\chapter{Implementation}

\section{Backup Server API Methods}
\label{sec:appendix-server-api}

The methods that the backup server's \verb!Client! API provides are as follows:

\begin{itemize}
    \item \verb!connected()! - the client has connected to the server
    \item \verb!disconnected()! - the client has disconnected from the server
    \item \verb!get_hostname()! - get the hostname of the client
    \item \verb!get_key()! - get the secret authentication key of the client
    \item \verb!get_paths()! - get the paths which are to be protected on the
        client
    \item \verb!get_exclusions()! - get the exclusions which are defined for
        the client
    \item \verb!get_present_state(path)! - get the present state of the
        client's file system from \verb!path! downwards, according to the
        catalog (including each item, and it's meta-data)
    \item \verb!is_excluded(path)! - check whether \verb!path! matches any
        exclusions (global or local) on the client
    \item \verb!create_item(path, type)! - create an item of type \verb!type!
        at \verb!path! in the client's file system catalog
    \item \verb!update_item(path, mtime, size, version_id)! - update the item
        at \verb!path! with the given meta data, and associate the new version
        with \verb!version_id! (for retrieving the file from the storage
        subsystem)
    \item \verb!delete_item(path)! - delete the item at \verb!path! in the
        client's file system catalog (marks the item as deleted in the catalog,
        rather than actually deleting it!)
    \item \verb!compare_attrs(old_attrs, new_attrs)! - calculates the
        difference between \verb!old_attrs! and \verb!new_attrs! as an integer
        signifying the amount of difference ($0$ meaning equal, $n$ meaning $n$
        attributes differ)
    \item \verb!backup_required(path, attrs)! - check whether the item at
        \verb!path! with attributes \verb!attrs! is in need of a backup (uses
        \verb!compare_attrs! to compare the attributes with those stored in the
        catalog)
    \item \verb!get_pending_restore_jobs()! - get a list of pending restore
        jobs for the client, in the format
        \verb!(job_id, source_path, version_id, destination_path)!
    \item \verb!restore_begin(job_id)! - notify the server that the restore job
        with ID \verb!job_id! has begun
    \item \verb!restore_complete(job_id)! - notify the server that the restore
        job with ID \verb!job_id! has completed
\end{itemize}

\section{Perspective Broker Methods}
\label{sec:appendix-server-pb-methods}

The backup server's perspective broker provides the following
remotely-accessible methods, which can be called by the client:

\begin{itemize}
    \item \verb!get_paths()! - get the paths which are to be backed up by this
        client
    \item \verb!get_present_state(path)! - get the present state of the
        client's file system from \verb!path! downwards, according to the
        catalog (including the name of each item, but \textbf{no} meta data)
    \item \verb!check_index(path, cur_index)! - check the given index
        (\verb!cur_index!) against that which exists in the catalog, retrieving
        the names of the files which require archival
    \item \verb!check_file(path, attrs)! - check if the file at \verb!path!
        with the attributes in \verb!attrs! requires backing up
    \item \verb!create_item(path, type)! - notify the server that the item at
        \verb!path! with type \verb!type! has been created on the file system
    \item \verb!delete_item(path)! - notify the server that the item at
        \verb!path! has been deleted from the file system (or moved to another
        location)
    \item \verb!put_file(path, mtime, size)! - transfer the file at \verb!path!
        (with the given \verb!mtime! and \verb!size!) to the archive; returns
        a remotely-referenceable \verb!PageCollector! object (see section
        \ref{sec:implementation-server-transfer}).
\end{itemize}
